Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-07-23 BayesianCubaturePaper2018 - Rev1b.tex
final(?)

% This is a general template file for the LaTeX package SVJour3
% for Springer journals. Original by Springer Heidelberg, 2010/09/16
%
% Use it as the basis for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
\RequirePackage{amsmath}
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{cite}
%
% insert here the call for the packages your document requires
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%\usepackage{latexsym}
% etc.
%
% Jag's 
\usepackage{tabu}
\usepackage{cancel}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
%\usepackage[caption=false]{subfig}
\usepackage{subcaption}
\usepackage{booktabs, mathtools}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
\usepackage[titletoc,title]{appendix}
%\usepackage{refcheck}  % points out unused labels, refs ...

@@

\begin{abstract}
Automatic cubatures approximate multidimensional integrals to user-specified error tolerances.  For high dimensional problems, it is difficult to adaptively change the sampling pattern, but one can automatically determine the
sample size, $n$, given a reasonable, fixed sampling pattern. We take this approach here using a Bayesian perspective.  We postulate that the integrand is an instance of a Gaussian stochastic process parameterized by a constant mean and a covariance kernel defined by a scale parameter times a parameterized function specifying how the integrand values at two different points in the domain are related.
These hyperparameters are inferred or integrated out using integrand values via one of three techniques:  empirical Bayes, full Bayes, or generalized cross-validation. The sample size, $n$, is increased until the half-width of the credible interval for the Bayesian posterior mean is no greater than the error tolerance.

@@

Users of cubature algorithms typically want the error to be no greater than their specified error tolerance, denoted by $\varepsilon$.  That is, they want
\begin{align}
\label{eqn:err_crit} 
\abs{\mu - \hmu} \leq \errtol .
\end{align}
Some stopping criteria for choosing $n$ are heuristic.  Other rigorous algorithms satisfying \eqref{eqn:err_crit}  require strong a priori assumptions about the integrand, such as an upper bound on its variance (for simple Monte Carlo) or total variation (for quasi-Monte Carlo).  We take a Bayesian approach by constructing a stopping criterion that is based on a credible interval.  We build upon the work of Briol et al.~\cite{BriEtal18a}, Diaconis~\cite{Dia88a}, O'Hagan~\cite{OHa91a}, Ritter~\cite{Rit00a}, Rasmussen and Ghahramani~\cite{RasGha03a}, and others.  Our algorithm is an example of \emph{probabilistic numerics}.

@@

We assume that the integrand, $f$, is an instance of a Gaussian stochastic process  i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$  \cite{BriEtal18a, Dia88a, OHa91a, Rit00a, RasGha03a}.  Specifically, $f$ is a real-valued random function with constant mean $m$ and covariance kernel $s^2C_\vtheta$:
\begin{gather*}
        m = \Ex[f(\vx)] \qquad \forall \vx \in \reals^d, \\
        \Ex\{[f(\vt) - m][f(\vx) -m]\} = s^2 C_\vtheta(\vt,\vx) \qquad \forall \vt, \vx \in \reals^d.
\end{gather*}
Here $s$ is a positive scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by the vector $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC_\vtheta^T = \mC_\vtheta,  \; \va^T \mC_\vtheta \va > 0,  \text{ where }  \mC_\vtheta = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \forall \va \ne 0, \;
 n\in \mathbb{N}, \; \text{ distinct }\vx_1, \ldots, \vx_n \in [0,1]^d.
\end{multline}
Procedures for estimating or integrating out the hyperparameters $m$, $s$, and $\vtheta$ are explained later in this section.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Empirical Bayes}  \label{sec:MLE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The empirical Bayes approach estimates the parameters, $m$, $s$, and $\vtheta$ via maximum likelihood estimation.  The log-likelihood function of the parameters given the function data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv_\vtheta(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det(\mC_\vtheta)) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$ and then with respect to $s$ yields the values given in Theorem \ref{thm:param}.
To obtain $\vtheta_\MLE$, we substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$, which leads directly to the optimization problem in \eqref{eqn:thetaMLE}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Full Bayes} \label{sec:fullBayes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Rather than use maximum likelihood to determine $m$ and $s$,  one can treat them as hyperparameters with a non-informative, conjugate prior, namely $\vrho_{m,s^2}(\xi, \lambda) \propto 1/\lambda$. We want to compute $\rho_{\mu|\vf}(z | \vy)$, the conditional posterior density of $\mu$ given the data $\vf = \vy$.  This may be expressed as 
\begin{multline*}
    \rho_{\mu|\vf}(z | \vy) 
= \int_{0}^\infty \int_{-\infty}^\infty 
\rho_{\mu | m, s^2, \vf}(z | \xi, \lambda , \vy) \\
\times  \rho_{m, s^2 | \vf}(\xi, \lambda | \vy)  \, \D \xi \D \lambda, 
\end{multline*}
where $\rho_{m, s^2 | \vf}$ is the posterior density of the hyperparameters given the function data. Bayes Theorem tells us that $\rho_{m, s^2 | \vf} \propto \rho_{\vf | m, s^2} \, \vrho_{m,s^2}$, so 
\begin{align}
\nonumber 
    \rho_{\mu|\vf}(z | \vy) 
& = \int_{0}^\infty \int_{-\infty}^\infty 
\rho_{\mu | m, s^2, \vf}(z | \xi, \lambda , \vy) \\
\nonumber
& \qquad \times  \rho_{\vf | m, s^2} (\vy | \xi, \lambda) \, \vrho_{m,s^2}(\xi,\lambda)  \, \D \xi \D \lambda \\
\nonumber
& \propto \left( 1 +  \frac{(z - \hmu_{\MLE})^2}{(n-1) \hsigma_{\textup{full}}^2} \right)^{-n/2},
\end{align}
where $\hsigma_{\textup{full}}^2$ is given in Theorem \ref{thm:param}, and the result above is derived in Appendix \ref{appendix:full_bayes}.

@@

As described above, the computational cost of Algorithm \ref{algorithm1} is the sum of the following:
\begin{itemize}
	\item $\Order\bigl(n\$(f) \bigr)$ for the function data, where $\$(f)$ is the computational cost of a single $f(\vx)$; $\$(f)$ may be large if it is the result of an expensive simulation; $\$(f)$ is typically proportional to $d$;

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval half-widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the function data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta,i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta, i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval half-widths simplify to
\begin{subequations}
\label{errSimple}
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\hspace{3cm} \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
\end{subequations}
In the formulas for the credible interval half-widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

We summarize the results of this section and the previous one in the theorem below.  In comparison to Algorithm \ref{algorithm1}, the second and third components of the computational cost of Algorithm  \ref{algorithm2} are substantially reduced.
\begin{theorem}
Let $C_\vtheta$ be any symmetric, positive definite,  shift-invariant covariance kernel of the form \eqref{eq:shInv}, where $K_\vtheta$ has period one in every variable.  Furthermore, let $K_\vtheta$ be scaled to satisfy \eqref{addAssump}.  When matched with rank-1 lattice data sites, $C_\vtheta$ must satisfy assumptions \eqref{fastcompAssump}.  The cubature, $\hmu$, is just the sample mean.  The \emph{fast Fourier transform} (FFT) can be used to expedite the estimates of $\vtheta$ in \eqref{fastTheta} and the credible interval half-widths \eqref{fastStoppingCriterions} so that Algorithm \ref{algorithm2} has a computational cost which is the sum of the following:
\begin{itemize}
	\item $\Order\bigl(n\$(f) \bigr)$ for the function data, where $\$(f)$ is the computational cost of a single $f(\vx)$;

@@

A noticeable observation from the plots in all three examples is how much the credible interval half-widths differ from the true error. The credible interval half-widths are not much smaller than $\varepsilon$.  The ratio of the true error to $\varepsilon$ ranges from around one to as small as $10^{-7}$.  In the latter case, the credible interval half-widths are quite conservative.  For option pricing example, this is less serious of an issue than for the multivariate Gaussian and Keister examples. A possible reason is that the latter integrands are significantly smoother than the covariance kernel assumed.  This is a matter for further investigation.

>>>>

% This is a general template file for the LaTeX package SVJour3
% for Springer journals. Original by Springer Heidelberg, 2010/09/16
%
% Use it as the basis for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
\RequirePackage{amsmath}
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amssymb}
%\usepackage{cite}
%
% insert here the call for the packages your document requires
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%\usepackage{latexsym}
% etc.
%
% Jag's 
\usepackage{tabu}
\usepackage{cancel}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
%\usepackage[caption=false]{subfig}
\usepackage{subcaption}
\usepackage{booktabs, mathtools}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
\usepackage[titletoc,title]{appendix}
%\usepackage{refcheck}  % points out unused labels, refs ...

@@

\begin{abstract}
Automatic cubatures approximate integrals to user-specified error tolerances.  For high dimensional problems, it is difficult to adaptively change the sampling pattern, but one can automatically determine the
sample size, $n$, given a reasonable, fixed sampling pattern. We take this approach here using a Bayesian perspective.  We postulate that the integrand is an instance of a Gaussian stochastic process parameterized by a constant mean and a covariance kernel defined by a scale parameter times a parameterized function specifying how the integrand values at two different points in the domain are related.
These hyperparameters are inferred or integrated out using integrand values via one of three techniques:  empirical Bayes, full Bayes, or generalized cross-validation. The sample size, $n$, is increased until the half-width of the credible interval for the Bayesian posterior mean is no greater than the error tolerance.

@@

Users of cubature algorithms typically want the error to be no greater than their specified error tolerance, denoted by $\varepsilon$.  That is, they want
\begin{align}
\label{eqn:err_crit} 
\abs{\mu - \hmu} \leq \errtol .
\end{align}
Some stopping criteria for choosing $n$ are heuristic.  Rigorous algorithms satisfying \eqref{eqn:err_crit}  typically require strong a priori assumptions about the integrand, such as an upper bound on its variance (for simple Monte Carlo) or total variation (for quasi-Monte Carlo).  We take a Bayesian approach by constructing a stopping criterion that is based on a credible interval.  We build upon the work of Briol et al.~\cite{BriEtal18a}, Diaconis~\cite{Dia88a}, O'Hagan~\cite{OHa91a}, Ritter~\cite{Rit00a}, Rasmussen and Ghahramani~\cite{RasGha03a}, and others.  Our algorithm is an example of \emph{probabilistic numerics}.

@@

We assume that the integrand, $f$, is an instance of a Gaussian stochastic process  i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$  \cite{BriEtal18a,Dia88a,OHa91a,Rit00a,RasGha03a}.  Specifically, $f$ is a real-valued random function with constant mean $m$ and covariance kernel $s^2C_\vtheta$:
\begin{gather*}
        m = \Ex[f(\vx)] \qquad \forall \vx \in \reals^d, \\
        \Ex\{[f(\vt) - m][f(\vx) -m]\} = s^2 C_\vtheta(\vt,\vx) \qquad \forall \vt, \vx \in \reals^d.
\end{gather*}
Here $s$ is a positive scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by the vector $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC_\vtheta^T = \mC_\vtheta,  \; \va^T \mC_\vtheta \va > 0,  \text{ where }  \mC_\vtheta = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \forall \va \ne 0, \;
 n\in \mathbb{N}, \; \text{ distinct }\vx_1, \ldots, \vx_n \in [0,1]^d.
\end{multline}
Procedures for estimating or integrating out the hyperparameters $m$, $s$, and $\vtheta$ are explained later in this section.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Empirical Bayes}  \label{sec:MLE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The empirical Bayes approach estimates the parameters, $m$, $s$, and $\vtheta$ via maximum likelihood estimation.  The log-likelihood function of the parameters given the integrand data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv_\vtheta(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det(\mC_\vtheta)) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$ and then with respect to $s$ yields the values given in Theorem \ref{thm:param}.
To obtain $\vtheta_\MLE$, we substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$, which leads directly to the optimization problem in \eqref{eqn:thetaMLE}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Full Bayes} \label{sec:fullBayes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Rather than use maximum likelihood to determine $m$ and $s$,  one can treat them as hyperparameters with a non-informative, conjugate prior, namely $\vrho_{m,s^2}(\xi, \lambda) \propto 1/\lambda$. We want to compute $\rho_{\mu|\vf}(z | \vy)$, the conditional posterior density of $\mu$ given the data $\vf = \vy$.  This may be expressed as 
\begin{multline*}
    \rho_{\mu|\vf}(z | \vy) 
= \int_{0}^\infty \int_{-\infty}^\infty 
\rho_{\mu | m, s^2, \vf}(z | \xi, \lambda , \vy) \\
\times  \rho_{m, s^2 | \vf}(\xi, \lambda | \vy)  \, \D \xi \D \lambda, 
\end{multline*}
where $\rho_{m, s^2 | \vf}$ is the posterior density of the hyperparameters given the integrand data. Bayes Theorem tells us that $\rho_{m, s^2 | \vf} \propto \rho_{\vf | m, s^2} \, \vrho_{m,s^2}$, so 
\begin{align}
\nonumber 
    \rho_{\mu|\vf}(z | \vy) 
& = \int_{0}^\infty \int_{-\infty}^\infty 
\rho_{\mu | m, s^2, \vf}(z | \xi, \lambda , \vy) \\
\nonumber
& \qquad \times  \rho_{\vf | m, s^2} (\vy | \xi, \lambda) \, \vrho_{m,s^2}(\xi,\lambda)  \, \D \xi \D \lambda \\
\nonumber
& \propto \left( 1 +  \frac{(z - \hmu_{\MLE})^2}{(n-1) \hsigma_{\textup{full}}^2} \right)^{-n/2},
\end{align}
where $\hsigma_{\textup{full}}^2$ is given in Theorem \ref{thm:param}, and the result above is derived in Appendix \ref{appendix:full_bayes}.

@@

As described above, the computational cost of Algorithm \ref{algorithm1} is the sum of the following:
\begin{itemize}
	\item $\Order\bigl(n\$(f) \bigr)$ for the integrand data, where $\$(f)$ is the computational cost of a single $f(\vx)$; $\$(f)$ may be large if it is the result of an expensive simulation; $\$(f)$ is typically proportional to $d$;

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval half-widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the integrand data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta,i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta, i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval half-widths simplify to
\begin{subequations}
\label{errSimple}
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\hspace{3cm} \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
\end{subequations}
In the formulas for the credible interval half-widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

We summarize the results of this section and the previous one in the theorem below.  In comparison to Algorithm \ref{algorithm1}, the second and third components of the computational cost of Algorithm  \ref{algorithm2} are substantially reduced.
\begin{theorem}
Let $C_\vtheta$ be any symmetric, positive definite,  shift-invariant covariance kernel of the form \eqref{eq:shInv}, where $K_\vtheta$ has period one in every variable.  Furthermore, let $K_\vtheta$ be scaled to satisfy \eqref{addAssump}.  When matched with rank-1 lattice data sites, $C_\vtheta$ must satisfy assumptions \eqref{fastcompAssump}.  The cubature, $\hmu$, is just the sample mean.  The \emph{fast Fourier transform} (FFT) can be used to expedite the estimates of $\vtheta$ in \eqref{fastTheta} and the credible interval half-widths \eqref{fastStoppingCriterions} so that Algorithm \ref{algorithm2} has a computational cost which is the sum of the following:
\begin{itemize}
	\item $\Order\bigl(n\$(f) \bigr)$ for the integrand data, where $\$(f)$ is the computational cost of a single $f(\vx)$;

@@

A noticeable observation from the plots in all three examples is how the ratio of the true error to the error tolerance varies from nearly one all the way down to $10^{-7}$.  Since the credible interval half-widths are not much smaller than $\varepsilon$, this means that the credible intervals are quite conservative in many cases.  For option pricing example, this is less of an issue than for the multivariate Gaussian and Keister examples. The reason that the credible intervals wildy overestimate the true error for the multivariate Gaussian and Keister examples may be that these integrands are significantly smoother than the assumed covariance kernel.  This is a matter for further investigation.
